### 2022-03-19 g-pldi-2022/texstyle.tex
pldi: cite software

\usepackage{mathpartir}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{positioning}

\usepackage[style=acmnumeric,backend=biber,natbib=true,datamodel=software]{biblatex}
\usepackage{software-biblatex}
\ExecuteBibliographyOptions{ halid=true, swhid=true, swlabels=true, vcs=true, license=true}

\settopmatter{printfolios=true,printccs=false,printacmref=false}

\renewcommand{\preDoc}{}

\input{def.tex}
\bibliography{bg}