### 2022-04-22 g-pldi-2022/texstyle.tex
pldi: fixes from Simone

\usepackage{mathpartir}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{positioning}

\usepackage[style=acmnumeric,backend=biber,natbib=true,datamodel=software]{biblatex}
\usepackage{software-biblatex}
\ExecuteBibliographyOptions{ halid=true, swhid=true, swlabels=true, vcs=true, license=true}

%\settopmatter{printfolios=true,printccs=false,printacmref=false}
%\overfullrule=1mm

\renewcommand{\preDoc}{}

\setcopyright{acmcopyright}
\acmPrice{15.00}
\acmDOI{10.1145/3519939.3523430}
\acmYear{2022}
\copyrightyear{2022}
\acmSubmissionID{pldi22main-p31-p}
\acmISBN{978-1-4503-9265-5/22/06}
\acmConference[PLDI '22]{Proceedings of the 43rd ACM SIGPLAN International Conference on Programming Language Design and Implementation}{June 13--17, 2022}{San Diego, CA, USA}
\acmBooktitle{Proceedings of the 43rd ACM SIGPLAN International Conference on Programming Language Design and Implementation (PLDI '22), June 13--17, 2022, San Diego, CA, USA}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011039.10011311</concept_id>
<concept_desc>Software and its engineering~Semantics</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011024.10011032</concept_id>
<concept_desc>Software and its engineering~Constraints</concept_desc>
<concept_significance>100</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
<concept_desc>Software and its engineering~Functional languages</concept_desc>
<concept_significance>100</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\input{def.tex}
\bibliography{bg}